---
title: "R Notebook"
output: html_notebook
---

```{text, Project Goal}
 A company recently introduced a new bidding type, “average bidding”, 
 as an alternative to its exisiting bidding type, called “maximum bidding”. 
 One of our clients has decided to test this new feature 
 and wants to conduct an A/B test to understand 
 if average bidding brings more conversions than maximum bidding.
 
 The A/B test ran for 1 month and the client now expects you 
 to analyze and present the results of this A/B test.
```


```{r, Import Block}
library(dplyr)
library(ggplot2)
library(reshape2)

average_bidding <- read.csv("control_group.csv", sep = ";", header = F)
maximum_bidding <- read.csv("test_group.csv", sep = ";", header = F)

```

```{r, Data Preview Block}

# View the first few rows of average_bidding data
head(average_bidding)

# View the first few rows of maximum_bidding data
head(maximum_bidding)

# Check the structure of average_bidding data
str(average_bidding)

# Check the structure of maximum_bidding data
str(maximum_bidding)

# Check if the number of rows is equal in both datasets
nrow_difference <- nrow(average_bidding) - nrow(maximum_bidding)
if (nrow_difference == 0) {
  print("Both datasets have the same number of rows.")
} else {
  print(paste0("Both datasets have a different number of rows. Difference: ", nrow_difference))
}

# Check if the number of columns is equal in both datasets
ncol_difference <- ncol(average_bidding) - ncol(maximum_bidding)
if (ncol_difference == 0) {
  print("Both datasets have the same number of columns.")
} else {
  print(paste0("Both datasets have a different number of columns. Difference: ", ncol_difference))
}

```

```{r, Quick View Block}

# View Head of Data
head(average_bidding)
head(maximum_bidding)

# View Structure of Data
str(average_bidding)
str(maximum_bidding)

# Check if both datasets have the same number of rows
print(paste("Number of rows in average_bidding: ", nrow(average_bidding)))
print(paste("Number of rows in maximum_bidding: ", nrow(maximum_bidding)))

# Check if both datasets have the same number of columns
print(paste("Number of columns in average_bidding: ", ncol(average_bidding)))
print(paste("Number of columns in maximum_bidding: ", ncol(maximum_bidding)))


```


```{r, Data Cleaning Block}

# Change the column names
column_names <- c("Campaign Name", "Date", "Spend (USD)",
                  "Impressions", "Reach", "Website Clicks", 
                  "Searches", "View Content", "Add to Cart", 
                  "Purchase")
colnames(maximum_bidding) <- column_names
colnames(average_bidding) <- column_names

# Remove the rows with column labels
average_bidding <- average_bidding[-1,] 
maximum_bidding <- maximum_bidding[-1,]       

# Change the data type of columns to integer
numeric_cols <- c("Spend (USD)", "Impressions", "Reach", "Website Clicks", 
                  "Searches", "View Content", "Add to Cart", "Purchase")
average_bidding[, numeric_cols] <- sapply(average_bidding[, numeric_cols], as.integer)
maximum_bidding[, numeric_cols] <- sapply(maximum_bidding[, numeric_cols], as.integer)

# Change the data type of the date column to Date
average_bidding$Date <- as.Date(average_bidding$Date, "%d.%m.%Y")
maximum_bidding$Date <- as.Date(maximum_bidding$Date, "%d.%m.%Y")

# Check for missing values
average_bidding_missing <- which(is.na(average_bidding) | average_bidding == "", arr.ind = TRUE)
average_bidding_missing_count <- sum(is.na(average_bidding))
maximum_bidding_missing <- which(is.na(maximum_bidding) | maximum_bidding == "", arr.ind = TRUE)
maximum_bidding_missing_count <- sum(is.na(maximum_bidding))

# Remove rows with missing values
average_bidding <- average_bidding[-5, ] # 7 Missing Values in this row 
maximum_bidding <- maximum_bidding[-5, ] # Removing to balance data

# Reset the row numbers
rownames(average_bidding) <- 1:nrow(average_bidding)
rownames(maximum_bidding) <- 1:nrow(maximum_bidding)

# Add Day of Week
average_bidding$day_of_week <- weekdays(average_bidding$Date)
maximum_bidding$day_of_week <- weekdays(maximum_bidding$Date)

```

```{r, Combining DataFrames for Futher Plotting}

# Melt the data into long format
columns_to_melt <- c("Spend (USD)", "Impressions", "Reach",
                     "Website Clicks", "Searches", "View Content",
                     "Add to Cart", "Purchase")

average_bidding_melt <- melt(average_bidding, id.vars = "Date", measure.vars = columns_to_melt)
maximum_bidding_melt <- melt(maximum_bidding, id.vars = "Date", measure.vars = columns_to_melt)

# Add a column indicating the group (average_bidding or maximum_bidding)
average_bidding_melt$group <- "average_bidding"
maximum_bidding_melt$group <- "maximum_bidding"

# Combine the two melted data frames into one data frame
combined_data <- rbind(average_bidding_melt, maximum_bidding_melt)

# Add day of week
combined_data$day_of_week <- weekdays(combined_data$Date)
average_bidding_melt$day_of_week <- weekdays(average_bidding_melt$Date)
maximum_bidding_melt$day_of_week <- weekdays(maximum_bidding_melt$Date)

combined_data

```

```{r, Comparing Single Variables Between Groups, One Plot At A Time}

# Plot Spend (USD) column
ggplot(data = average_bidding) + 
  geom_line(aes(x = Date, y = `Spend (USD)`, color = "Average Bidding")) +
  geom_line(data = maximum_bidding, aes(x = Date, y = `Spend (USD)`, color = "Maximum Bidding")) +
  ggtitle("Spend (USD) Over Time") + 
  xlab("Date") + 
  ylab("Spend (USD)") + 
  scale_color_manual(values = c("Average Bidding" = "blue", "Maximum Bidding" = "red"))

# Plot Impressions column
ggplot(data = average_bidding) + 
  geom_line(aes(x = Date, y = Impressions, color = "Average Bidding")) +
  geom_line(data = maximum_bidding, aes(x = Date, y = Impressions, color = "Maximum Bidding")) +
  ggtitle("Impressions Over Time") + 
  xlab("Date") + 
  ylab("Impressions") + 
  scale_color_manual(values = c("Average Bidding" = "blue", "Maximum Bidding" = "red"))

# Plot Reach column
ggplot(data = average_bidding) + 
  geom_line(aes(x = Date, y = Reach, color = "Average Bidding")) +
  geom_line(data = maximum_bidding, aes(x = Date, y = Reach, color = "Maximum Bidding")) +
  ggtitle("Reach Over Time") + 
  xlab("Date") + 
  ylab("Reach") + 
  scale_color_manual(values = c("Average Bidding" = "blue", "Maximum Bidding" = "red"))

# Plot Website Clicks column
ggplot(data = average_bidding) + 
  geom_line(aes(x = Date, y = `Website Clicks`, color = "Average Bidding")) +
  geom_line(data = maximum_bidding, aes(x = Date, y = `Website Clicks`, color = "Maximum Bidding")) +
  ggtitle("Website Clicks Over Time") + 
  xlab("Date") + 
  ylab("Website Clicks") + 
  scale_color_manual(values = c("Average Bidding" = "blue", "Maximum Bidding" = "red"))


```



```{r, Comparing All Variables of Both Groups in By Day of Week to See the Performance Differences by Day}

# Varible By Day of The Week

average_bidding_melt %>% 
  group_by(day_of_week, variable) %>% 
  summarise(sum_of_values = sum(value)) %>% 
  ggplot(aes(x = day_of_week, y = sum_of_values, fill = variable)) + 
  geom_col() + 
  labs(x = "Day of the week", y = "Sum of values") + 
  ggtitle("Sum of values by day of the week and variable") + 
  facet_wrap(~ variable, ncol = 1,scales = "free_y")

# Percent of Total by Day of Week

average_bidding_melt %>%
  group_by(day_of_week, variable) %>%
  summarise(sum_of_values = sum(value)) %>%
  group_by(variable) %>%
  mutate(percent_of_total = sum_of_values / sum(sum_of_values) * 100) %>%
  ggplot(aes(x = day_of_week, y = percent_of_total, fill = variable)) +
  geom_col() +
  geom_text(aes(label = sprintf("%1.1f%%", percent_of_total)), position = position_fill(vjust = 0.75)) +
  labs(x = "Day of the week", y = "Percent of total") +
  ggtitle("Percent of total by day of the week and variable") +
  facet_wrap(~ variable, ncol = 1, scales = "free_y")

# As a ratio of spend: can adjust for other ratios

combined_data %>% 
  group_by(day_of_week, variable) %>% 
  summarise(sum_of_values = sum(value)) %>% 
  mutate(ratio = sum(sum_of_values[variable == "Spend (USD)"]) / sum_of_values) %>% 
  ggplot(aes(x = day_of_week, y = ratio, fill = variable)) + 
  geom_col() + 
  geom_text(aes(label = round(ratio, 2)), position = position_dodge(width = 0.9), vjust = -0.25) +
  labs(x = "Day of the week", y = "Ratio") + 
  ggtitle("Ratio of each variable to spend by day of the week") + 
  facet_wrap(~ variable, ncol = 1,scales = "free_y")

# We can analyize these groups to see how much our costs are per action per day of the week
# 

```

```{r, EDA for Plotting by Day of Week}

# Spend By Day of The Week as Sum
combined_data %>%
  filter(variable == "Spend (USD)") %>%
  group_by(day_of_week) %>%
  summarise(sum_of_values = sum(value)) %>% 
  print(n = 200)

# Spend By Day of The Week as Percentage
combined_data %>%
  filter(variable == "Spend (USD)") %>%
  group_by(day_of_week) %>%
  summarise(sum_of_values = sum(value)) %>% 
  mutate(percent_of_total = sum_of_values / sum(sum_of_values) * 100)


# All Variables by Day of The Week with Percentage
combined_data %>%
  group_by(day_of_week, variable) %>%
  summarise(sum_of_values = sum(value)) %>%
  group_by(variable) %>%
  mutate(percent_of_total = sum_of_values / sum(sum_of_values) * 100)

# Ratio of Spend to Every Variable of the group
combined_data %>%
  group_by(day_of_week, variable) %>%
  summarise(sum_of_values = sum(value)) %>%
  mutate(ratio = ifelse(variable != "Spend (USD)", sum_of_values / sum(sum_of_values[variable == "Spend (USD)"]), NA))


```

```{r, Facet Wrap Comparison of All Variable Daily Values to Compare Both Groups}

# Facet bar plot
ggplot(combined_data, aes(x = Date, y = value, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~variable, scales = "free_y") +
  ggtitle("Bar Plot Comparison") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, hjust = 1))

# Facet line plot
ggplot(combined_data, aes(x = Date, y = value, color = group)) +
  geom_line(aes(group = group)) +
  facet_wrap(~variable, scales = "free_y") +
  ggtitle("Line Plot Comparison") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Stacked Plots
ggplot(combined_data, aes(x = Date, y = value, fill = group)) +
  geom_area(position = "stack") +
  ggtitle("Stacked Plot Comparsion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~variable, scales = "free_y")

# Can Use This Analysis to Easily Spot Outliers

```

```{r, Monthly Data EDA and Graphing}

# Aggregate data to get sum of each column
monthly_data <- combined_data %>%
  group_by(group,variable) %>%
  summarize(sum = sum(value))

# Cost per Variable
monthly_data_ratio <- combined_data %>%
  group_by(group, variable) %>%
  summarise(sum = sum(value)) %>%
  mutate(ratio = ifelse(variable != "Spend (USD)", sum(sum[variable == "Spend (USD)"]), NA) / sum) %>%
  filter(variable != "Spend (USD)")
  monthly_data_ratio
  
# Purchases per Variable
monthly_data_ratio_purchase <- combined_data %>%
  group_by(group, variable) %>%
  summarise(sum = sum(value)) %>%
  mutate(ratio = sum / ifelse(variable != "Purchase", sum(sum[variable == "Purchase"]), NA)) %>%
  filter(variable != "Purchase")
  monthly_data_ratio_purchase


# Validating the Sum of Values
average_bidding %>%
  summarize(
    Spend = sum(`Spend (USD)`),
    Impressions = sum(Impressions),
    Reach = sum(Reach),
    Website_Clicks = sum(`Website Clicks`),
    Searches = sum(Searches),
    View_Content = sum(`View Content`),
    Add_to_Cart = sum(`Add to Cart`),
    Purchase = sum(`Purchase`)
  )
# average_bidding 66818	3177233	2576503	154303	64418	56370	37700	15161
# maximum_bidding 74595	2123249	1509609	175107	70463	54882	25490	14869

# Sum of Each Variable for the Period
ggplot(monthly_data, aes(x = group, y = sum, fill = group)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = round(sum)), position = position_stack(vjust = 0.5)) +
  facet_wrap(~variable, scales = "free_y",ncol=2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Cost of Each Variable
ggplot(monthly_data_ratio, aes(x = group, y = ratio, fill = variable)) +
  geom_col() +
  geom_text(aes(label = round(ratio, 2)), position = position_dodge(width = 0.9), vjust = -0.25) +
  labs(x = "Group", y = "Ratio to Spend") +
  ggtitle("Ratio of each variable to Spend by Group") +
  facet_wrap(~ variable, ncol = 1,scales = "free_y") +
  theme(plot.title = element_text(hjust = 0.5))

# Purchase to Variable Ratio
ggplot(monthly_data_ratio_purchase, aes(x = group, y = ratio, fill = variable)) +
  geom_col() +
  geom_text(aes(label = round(ratio, 2)), position = position_dodge(width = 0.9), vjust = -0.25) +
  labs(x = "Group", y = "Ratio to Spend") +
  ggtitle("Ratio of each variable to Spend by Group") +
  facet_wrap(~ variable, ncol = 1,scales = "free_y") +
  theme(plot.title = element_text(hjust = 0.5))


# Needed less impressions, clicks, and add-to-carts to get to a purchase with Maximum Bidding
# Why was the spend higher, impressions and reach lower, Clicks, Searches Lower, with better Add-to-cart to Purchase Ratios?
```

```{r}
agg_data$ratio <- 0

# Calculate the ratio of each variable to spend for the control group
agg_data$ratio[agg_data$variable != "Spend (USD)" & agg_data$group == "control"] <-  agg_data$sum[agg_data$variable == "Spend (USD)" & agg_data$group == "control"] / agg_data$sum[agg_data$variable != "Spend (USD)" & agg_data$group == "control"]

# Calculate the ratio of each variable to spend for the test group
agg_data$ratio[agg_data$variable != "Spend (USD)" & agg_data$group == "test"] <- agg_data$sum[agg_data$variable == "Spend (USD)" & agg_data$group == "test"] / agg_data$sum[agg_data$variable != "Spend (USD)" & agg_data$group == "test"]

ggplot(subset(agg_data, variable != "Spend (USD)"), aes(x = group, y = ratio, fill = group)) +
  geom_bar(stat="identity", position="stack") +
  geom_text(aes(label=round(ratio, 2)), position=position_stack(vjust=0.5)) +
  labs(x="Group", y="Ratio of Spend to Variable", fill="Group") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  facet_wrap(~variable, ncol=1, scales="free_y")

```


# -------------------------------------------- Aggregated Data for the month


# -------------------------------------------- Monthly Sales to Variable Ratio : Select the Variable for a Ratio

# A statistical comparison to see if the values are significantly different basd on the number of obersvations

# ------------ spend / purchase, t test, 2 sample t test
combined_data spend/purchase, spend/purchase b, visualize in overlaying distribution of both
t test value

combined_data %>%
  filter(variable %in% c("Spend (USD)", "Purchase"))

combined_data %>%
  filter(variable %in% c("Spend (USD)", "Purchase")) %>%
  pivot_wider(names_from = c(variable, group), values_from = value) %>%
  select(-day_of_week) %>%
  mutate(control =  `Spend (USD)_control` / `Purchase_control`,
         test =  `Spend (USD)_test` / `Purchase_test`)


  ggplot(aes(x = control, y = test)) +
  geom_point() +
  labs(x = "Control", y = "Test") +
  ggtitle("Control vs Test")

combined_data %>%
  filter(variable %in% c("Spend (USD)", "Purchase")) %>%
  pivot_wider(names_from = c(variable, group), values_from = value) %>%
  mutate(control =  `Spend (USD)_control` / `Purchase_control`,
         test =  `Spend (USD)_test` / `Purchase_test`) %>%
  select(control,test)  %>%
  print(n=50)
    
combined_data %>%
  filter(variable %in% c("Spend (USD)", "Purchase")) %>%
  pivot_wider(names_from = c(variable, group), values_from = value) %>%
  select(-day_of_week) %>%
  mutate(control =  `Spend (USD)_control` / `Purchase_control`,
         test =  `Spend (USD)_test` / `Purchase_test`) %>% 
  ggplot(aes(x = control)) +
  geom_density(aes(fill = "control"), alpha = 0.5) +
  geom_density(aes(x = test, fill = "test"), alpha = 0.5) +
  scale_fill_manual(values = c("control" = "red", "test" = "green")) +
  ggtitle("Density Plot Overlay of Control vs Test")

# Control test has lower spend per purchase

# Statistical analysis

control <- c(3.69, 3.44, 6.30, 5.71, 4.04, 5.10, 4.11, 5.61, 2.93, 5.24, 2.92, 3.52, 2.38, 7.58, 4.62, 9.81, 7.72, 4.79, 8.95, 2.25, 7.59, 3.26, 3.23, 5.08, 8.90, 3.41, 3.77, 7.11, 3.47)
test <- c(11.8, 3.75, 4.09, 7.97, 5.04, 3.19, 6.77, 3.14, 10.1, 3.62, 3.99, 2.43, 7.46, 6.32, 6.32, 8.27, 7.70, 5.13, 3.72, 12.7, 10.5, 5.09, 7.72, 4.50, 8.14, 3.78, 3.12, 4.14, 3.46)

t.test(control, test)

wilcox.test(control, test, alternative = "two.sided")

# more days of analysis may be needed
# how many more days would be needed --> power analysis


# analyize for outliers and anamolies

```{r, Descriptive Statistics Block (undeveloped)}

# Set Column Names
col_names <- c("Spend (USD)", "Impressions", "Reach", "Website Clicks", "Searches", "View Content", "Add to Cart", "Purchase")

# Average Bidding Group
means_avg_bidding <- colMeans(average_bidding[, col_names], na.rm = TRUE)
stddevs_avg_bidding <- apply(average_bidding[, col_names], 2, sd, na.rm = TRUE)
q1_avg_bidding <- apply(average_bidding[, col_names], 2, quantile, probs = c(0.25), na.rm = TRUE)
q3_avg_bidding <- apply(average_bidding[, col_names], 2, quantile, probs = c(0.75), na.rm = TRUE)
iqr_avg_bidding <- q3_avg_bidding - q1_avg_bidding

desc_stats_avg_bidding <- data.frame(mean = means_avg_bidding, stddev = stddevs_avg_bidding,
                                     q1 = q1_avg_bidding, q3 = q3_avg_bidding, iqr = iqr_avg_bidding)

# Maximum Bidding Group
means_max_bidding <- colMeans(maximum_bidding[, col_names], na.rm = TRUE)
stddevs_max_bidding <- apply(maximum_bidding[, col_names], 2, sd, na.rm = TRUE)
q1_max_bidding <- apply(maximum_bidding[, col_names], 2, quantile, probs = c(0.25), na.rm = TRUE)
q3_max_bidding <- apply(maximum_bidding[, col_names], 2, quantile, probs = c(0.75), na.rm = TRUE)
iqr_max_bidding <- q3_max_bidding - q1_max_bidding

desc_stats_max_bidding <- data.frame(mean = means_max_bidding, stddev = stddevs_max_bidding,
                                     q1 = q1_max_bidding, q3 = q3_max_bidding, iqr = iqr_max_bidding)

```

